What Can Stack Traces Tell us about Exceptions in Java?

### 2014-07-07 android-stacks.tex
A new take on the introduction

@@

Bug issues have called the attention of research community and have been the target of several 
researches [][][]. A subtype of a bug issue are the ones related to system 
crashes, also called crash reports such reports can be automatically generated 
as the system crashes or can be manually reported by developers/users [complete]...

One of the most important elements of a crash report genereted eigther manually
or automatically is the failure stack trace [ref-paper], which tipically consists 
of an ordered list of methods that were active on the call stack before the failure occurred.

Such failure stack traces are a useful source of 
information about system crashes, and are often used to
support developers in debugging~\cite{schroter2010stack}. Moreover, information
from stack traces combined with crash and bug reports can be used to support bug
classification and clustering~\cite{wang2013improving, kim2011crash,
dhaliwal2011classifying}, fault-proneness prediction
models~\cite{kim2013predicting} and even automated bug
fixing~\cite{sinha2009fault} tools.

The exception handling mechanism~\cite{goodenough1975exception},
 is one of the most used schemes for detecting and recovering from such 
failures. Exceptions may arrise as a consequence of faults in the application
itself (e.g., access of null references), noisy user inputs or faults in
underlying middleware or hardware.

Although moder application have to deal with a similar and increasing number of 
source of failures, existing research has focused on the analysis of a 
single system at a time.

@@

Hence, in such context where we have access to wealth of information available
on open-source repositories such as GitHub, some questions arise: What are the common
characteristics of Github crash reports?

@@

....... OLD VERSION ......

@@

Hence, in Java an exception can be thrown in one of the following
circumstances~\cite{gosling2000java}: (i) explicitly thrown when a throw statement is executed; 
(ii) implicitly thrown by the JVM when the evaluation of an expression
 violates the normal semantics of language (e.g., out-of-bounds array index,
division-by-zero, access to a null reference); or (iii) implicitly thrown by
the JVM due to an internal error or resource limitation (e.g.,
OutofMemoryError).

@@

When an exception is throw it causes a transfer of control from the point
where the exception occurred to a point that can be specified by the programmer
(exception handler) - in Java it is represented by the try-catch block .

A common way of  propagating exceptions in Java programs, is the exception wrapping (also known as exception chaining), which allows one exception to be wrapped in another exception and re-thrown. 
Figure~\ref{fig:wrapping} presents a stack trace associated which illustrates an
exception wrapping. The bottom part of the stack trace is the \emph{root cause}, which indicates the
first reason for the error to be thrown (in this case, the computer run out of
memory). The top part of the stack trace indicates the location of the exception
manifestation (which will call \emph{top level exception} along this paper). The
execution flow  between the root cause and the exception manifestation may
include other intermediate exception wrappings. In all levels, the exception
\emph{signaler}, is the method that threw the exception, represented on the
stack trace as the first method call below the exception declaration.

@@

This work describes an exploratory study which was based on a sequential
mixed-methods approach~\cite{ivankova2006using} for collecting, analyzing, and
integrating both quantitative and qualitative evidence along the stages of the
research process.

The main goal of this study was to gain a better understanding
of the use of exceptions in Java programs based mainly on the information
available on exception stack traces embedded on Github issues.
The exceptions available on exception stack trace are not limited to the exceptions 
expliciltly thrown by the Java developer but also include
the exceptions implicitly thrown by the JVM during the program execution,
due to a programming error (e.g., out-of-bounds array index, division-by-zero, 
access to a null reference) or an internal error or resource limitation (e.g.,
OutofMemoryError).

 More specifically, we aimed at answering
the following research questions:

@@

  \item[RQ1] What are the common characteristics of such stack traces?

  \item[RQ2] What patterns of exception (mis)-use emerge from the stack trace
    analysis?

@@

For RQ1 and RQ2, we explore the domain quantitatively and highlight interesting cases by 
exploring cases qualitatively. To support this investigation, source code and bytecode 
analysis were used in combination with manual inspection to to everage the  understanding of stack traces.

As mentioned before, this study amined at mining information available on GitHub issues,
more specifically, extracting and distilling the stack traces embedded on GItHub issues.

Issues on Github are different from issues on dedicated bug tracking tools such as 
Bugzilla and Jira. The most important difference is that there are no predefined fields
  (e.g. severity and priority). Instead, Github uses a more open ended tagging system, on which
repositories are offered a pre-defined set of labels, but repository owners can modify 
them at will. Therefore, an issue may have none ore an arbitrary set of labels depending 
on which repository it was created.

Therefore, this study is based on the assumption if an stack trace information embedded
 in an issue, it contain relevant information concerning the exception use -  regardless the
 lables associated to the issue.

Moreover, issues and pull requests are dual on Github; all pull requests have a corresponding 
``backing'' issue which
are automatically generated. Therefore, we excluded such automatically generated
issues from our analysis. Finally, the Github API allows the automated
generation of issues for specific repositories, which automated tools often use
to report crashes. In some cases, this led to high number of issues that
included stack traces. We identified and filtered those cases out (e.g.,
the~\textsf{pullwifi} project was responsible for almost 50\% of all Android issues in our dataset).

This study used the dataset provided by the GHTorrent project~\cite{Gousi13}, an off-line mirror of the data 
offered through the Github API.  GHTorrent has been collecting data from Git since 
February 2012. Up to Dec 2013, and when we queried GHTorrent, it included 332,864 non-fork Java
repositories \footnote {A Java project is specified in GitHub with a
specific label.}  from which only 44,323 have contained at least issue report in their lifetime. From this set only 16,837 were watched by external users. To ensure that our work targets projects that are openly used by
users other than their developers, we focused our analysis on just those projects.

From this set of 16.836 Java projects which overall contained 356.057 issues, we retrieved the metadata related to each Java project (e.g., repository's names and short descriptions), and the content of every issue and issue comment and stored in a relational database.

Using the heuristics presented in Section~\ref{sec:android}, we selected 482 repositories featuring Android projects on which we perform manual inspection was performed.

\subsection{Study Context}

Before investigating the research questions listed above we analyzed the prevalence of stack traces 
on issues created on Github. Prevalence or prevalence proportion, a term often used in epidemiology,
represents the proportion of a population found to have a particular condition
(e.g. a disease) over the total number of the studied population. In our work, we used
this term to evaluate a condition (i.e., presence of a stack trace) over issues
reported on GitHub repositories.

\subsection{The Mining Process}
\label{sec:miningproc}

Figure~\ref{fig:overviewfig} presents an overview of the mixed-methods approach
conducted to answer these questions.

@@

@@

@@

TO DISCUSS: How can we benefit from such cross-project crash analysis?

>>>>

\subsection{Georgios's Introduction}

Modern applications have to cope with an increasing number of abnormal
computation states that arise as a consequence of faults in the application
itself (e.g., access of null references), noisy user inputs or faults in
underlying middleware or hardware. The exception handling
mechanism~\cite{goodenough1975exception} is one of the most used schemes for
detecting and recovering from such exceptional conditions. When an exception is
caught by the program's exception handling code, the program has the opportunity
to assess the conditions that caused the fault and either recover the
computation or abort. An exception that is not caught by an exception handler
leads to a program crash. In most programming languages with exceptions, a crash
usually causes debugging information to emited by the language runtime to help
the programmer identify and fix the reason of the crash. The debugging
information is typical a stack trace, which includes an explanatory message and
the execution stack frame when the exception occurred. In Java, stack trace
consists of an ordered list of methods that were active on the call stack before
the exception occurred.

Although the exception handling mechanism has been the target of several studies
(e.g.~\cite{miller1997issues,Robil00,shah2010understanding,
garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
the exception handling code is often generally poorly understood and the least
tested part of software systems~\cite{coelho2011unveiling}. Similarily to source
code, current works focus their exception handling analysis on either static
(source code-based) or dynamic (runtime-based) analysis methods. Neither method
can explore the full execution space of a program and therefore they both fail
to account for exceptions caused by rutime issues (e.g. resource depletion) or
dynamically loaded code. Exception stack traces, when available, enable the
post-mortem analysis of exception handling behaviours of programs and
concequently the evaluation of exception handling mechanisms in programming
languages.

The availability of exception stack traces until recently was scarce, while the
available data were only focused on few large scale projects. In the recent
past, many open source software projects have their source code and issue
repositories to Github. The concentration of projects and the existence of
projects that mine such information, such as GHTorrent~\cite{Gousi13}, made full
population studies possible.

In this work, we perform the first post mortem analysis of exception handling in
Java.  Using a custom tool called ExceptionMiner, which we develop specifically
for this study, we mine stack traces from issues across all Java projects on
GitHub. Overall, we analyzed 356,057 issues from 16,836 projects, from which
28,800 stack traces were extracted and used as a source of information for our
analysis. The stack trace analysis was augmented by additional information
extracted using byte code and source code analysis done on a carefully selected
subset of our initial sample (482 Android projects).
To guide this exploration we compiled general guidelines on how to use
exceptions proposed by Gosling~\cite{gosling2000java},
Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then
we investigate what stack traces can tell us about the adherence to such
practices in Java programs.

Some outcomes were consistently detected through
our large scale analysis of exception stack traces in the wild, such as:

The contributions of this study are as follows:
\begin{itemize}

%This paper is organized as follows. Section 2 presents a
%background on exception handling mechanism. Section 3 presents the study design.
%Section 4 reports study findings. Section 5 provides further discussions and insights.
%Section 6 presents the threats to validity associated to this study. Finally Section
%7 describes the related work, and Section 8 presents our conclusions and
%directions for future work.

\subsection{Remaining Introductions :-)}

@@

Bug issues have called the attention of researhc community and  have been the target of several 
researches [][][]. A subtype of a bug issue are the ones related to system 
crashes, also called crash reports such reports can be automatically generated 
as the system crashes or can be manually reported by developers/users [complete]...

In [ref] defined a set of characteristics a good bug issue should have. They showed that among 
the most important relevant information are: (1) steps to reproduce the bug; 
(2) and the failure stack trace.

A typical stack trace consists of an ordered list of methods that
were active on the call stack before the failure occurred. Stack traces
are a useful source of information about system crashes, and are often used to
support developers in debugging~\cite{schroter2010stack}. Moreover, information
from stack traces combined with crash and bug reports can be used to support bug
classification and clustering~\cite{wang2013improving, kim2011crash,
dhaliwal2011classifying}, fault-proneness prediction
models~\cite{kim2013predicting} and even automated bug
fixing~\cite{sinha2009fault} tools.

We could observe that existing research has focused on the analysis of a 
single system at a time. Moreover, the search facilities of repositories are very limited, also
enabling the analysis of a single project.

Modern applications have to cope with similar and increasing number of abnormal
computation states that arise as a consequence of faults in the application
itself (e.g., access of null references), noisy user inputs or faults in
underlying middleware or hardware. The exception handling mechanism~\cite{goodenough1975exception} is one of the most used schemes for detecting and recovering from such exceptional conditions.

@@

Hence, in such context where we have access to wealth of information available
on open-source repositories such as GitHub, some questions arise: What are the common
characteristics of Github crash reports? How can we benefit from such cross-project crash analysis?

@@

-----
Along the recent years we have witnessed a rapid increase in 
the number of social applications on the Web. Applications which 
allow their users to track the activities of a large number 
of others regardless of his/her location and affiliation.

This tendency can also be seen on the software development context 
where Github emerged and became an extensively used social coding 
platform. Github alows developers to create and maintain opensource 
projects relying  on Git distributed version control, and provides a set 
of tools which allow the information about developers and their activities
 to be visible within and across open souce projects.

Among those tools there is the Github issue tracking system, which makes
issues related to new, features, bugs and general comments publicly available.

-----

@@

%\subsubsection{Exception Throwing and Handling}

In Java an exception can be thrown in one of the following
circumstances~\cite{gosling2000java}: (i) explicitly thrown when a throw statement is executed; 
(ii) implicitly thrown by the JVM when the evaluation of an expression
 violates the normal semantics of language (e.g., out-of-bounds array index,
division-by-zero, access to a null reference); or (iii) implicitly thrown by
the JVM due to an internal error or resource limitation (e.g.,
OutofMemoryError).

@@

%When an exception is signaled it causes a transfer of control from the point
%where the exception occurred to a point that can be specified by the programmer
%(exception handler) - in Java it is represented by the try-catch block .
%The search for the handler to deal with a raised
%exception occurs along the dynamic invocation chain. 
%If no handler is found in the chain, the default handler associated to the 
%Thread or ThreadGroup is invoked, which usually terminates the system 
%and prints the exception stack trace.

%\subsubsection{Exception Chaining} Exception Wrapping 
%This stack trace was extracted from an issue of an Android application in
%Github.

The Exception wrapping (also known as exception chaining) is a common way of 
propagating exceptions in Java programs, 
which allows one exception to be wrapped in another exception and re-thrown. 
Figure~\ref{fig:wrapping} presents a stack trace associated which illustrates an
exception wrapping. The bottom part of the stack trace is the \emph{root cause}, which indicates the
first reason for the error to be thrown (in this case, the computer run out of
memory). The top part of the stack trace indicates the location of the exception
manifestation (which will call \emph{top level exception} along this paper). The
execution flow  between the root cause and the exception manifestation may
include other intermediate exception wrappings. In all levels, the exception
\emph{signaler}, is the method that threw the exception, represented on the
stack trace as the first method call below the exception declaration.

@@

This work describes an exploratory study which was based on a sequential
mixed-methods approach~\cite{ivankova2006using} for collecting, analyzing, and
integrating both quantitative and qualitative evidence along the stages of the
research process. The main goal of this study was to gain a better understanding
of the use of exceptions in Java programs based mainly on the information
available on exception stack traces. More specifically, we aimed at answering
the following research questions:

@@

  \item[RQ1] What is the prevalence of stack traces on  reported issues?
    %\item[RQ2] What are the most common exceptions?

  \item[RQ2] What are the common characteristics of such stack traces?

  \item[RQ3] What patterns of exception (mis)-use emerge from the stack trace
    analysis?

@@

For RQ1 and RQ2, we explore the domain quantitatively, and use their answers as
a basis to investigate the third research question. RQ3 aims at examining whether
exception (mis)-uses could be revealed through stack trace analysis. To
support this investigation, source code and bytecode analysis was used to
leverage the  understanding of stack traces. 
%For RQ3 we investigate
%quantitatively and highlight interesting cases by exploring cases qualitatively.
Figure~\ref{fig:overviewfig} presents an overview of the mixed-methods approach
conducted to answer these questions.

@@

\noindent\emph{Kinds of Analyzed Issues.} Issues on Github are different from
issues on dedicated bug tracking tools such as Bugzilla and Jira. The most
important difference is that there are no predefined fields for bug reporters to
select from (e.g. severity and priority); Github uses instead a more open ended
tagging system. Repositories are offered a pre-defined set of labels, but
repository owners can modify them at will. An issue can have an arbitrary set of
labels attached. Hence, this study does not restrict the analysis of stack
traces defined on defect-issues, based on the assumption that regardless of the
type of issue where its is defined the stack trace information contain relevant
information concerning the exception use. Moreover, issues and pull requests are
dual on Github; all pull requests have a corresponding ``backing'' issue which
are automatically generated. Therefore, we excluded pull request based
issues from our analysis. Finally, the Github API allows the automated
generation of issues for specific repositories, which automated tools often use
to report crashes. In some cases, this led to high number of issues that
included stack traces. We identified and filtered those cases out (e.g.,
the~\textsf{pullwifi} project was responsible for almost 50\% of all Android issues in our dataset).

@@

Prevalence or prevalence proportion, a term often used in epidemiology,
represents the proportion of a population found to have a particular condition
(e.g. a disease) over the total number of the studied population. In our work, we used
this term to evaluate a condition (i.e., presence of a stack trace) over issues
reported on GitHub repositories.

@@
